---
title: "PresentHW"
author: "Samuel Kigamba"
date: "Sep 1, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## On Your Own

In the previous few pages, you recreated some of the displays and preliminary 
analysis of Arbuthnot's baptism data. Your assignment involves repeating these 
steps, but for present day birth records in the United States.  Load up the 
present day data with the following command.

```{r load-present-data, eval=TRUE}
source("more/present.R")
```

```{r view-present-data, eval=TRUE}
head(present, 5)
tail(present, 5)
```
```{r load-abrbuthnot-data, eval=TRUE}
source("more/arbuthnot.R")
```

```{r view-abrbuthnot-data, eval=TRUE}
head(arbuthnot, 5)
tail(arbuthnot, 5)
```
What years are included in this data set?
```{r view-years-present, eval=TRUE}
present$year
```
The year is 1940-2002

What are the dimensions of the data frame?
```{r view-dimensions-present, eval=TRUE}
dim(present)
```
63 rows and 3 columns
    
What are the variable or column names?
```{r view-column-names-present, eval=TRUE}
names(present)
```
"year"  "boys"  "girls"

How do these counts compare to Arbuthnots? Are they on a similar scale?

both have 3 columns, "year"  "boys"  "girls".
```{r view-column-names-arbuthnot, eval=TRUE}
names(arbuthnot)
```
```{r view-compare-column-names-present, eval=TRUE}
names(present)
```

Arbuthnot has 82 rows while present has 63,
```{r view-last-row-present-data, eval=TRUE}
tail(present, 1)
```
```{r view-last-row-abrbuthnot-data, eval=TRUE}
tail(arbuthnot, 1)
```

Make a plot that displays the boy-to-girl ratio for every year in the data set. 
    
```{r plot-ratio-of-boys-to-girls-present, eval=TRUE}
plot(present$year, present$boys / present$girls, type = "l")
```

What do you see? The ratio of boys to girls is higher although reducing for all the years.


Does Arbuthnots observation about boys being born in greater proportion than girls hold up in the U.S.? yes it does.
    
Include the plot in your response.

```{r plot-ratio-of-boys-to-girls-arbuthnot, eval=TRUE}
plot(arbuthnot$year, arbuthnot$boys / arbuthnot$girls, type = "l")
```


In what year did we see the most total number of births in the U.S.? Year 1961 total population of 4,268,326.
```{r max-population-year-present, eval=TRUE}
present[present$boys + present$girls == max(present$boys + present$girls),]
```


You can
    refer to the help files or the R reference card 
    [http://cran.r-project.org/doc/contrib/Short-refcard.pdf](http://cran.r-project.org/doc/contrib/Short-refcard.pdf)
    to find helpful commands.

These data come from a report by the Centers for Disease Control 
[http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf](http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf). 
Check it out if you would like to read more about an analysis of sex ratios at 
birth in the United States.

That was a short introduction to R and RStudio, but we will provide you with more
functions and a more complete sense of the language as the course progresses. 
Feel free to browse around the websites for [R](http://www.r-project.org) and 
[RStudio](http://rstudio.org) if you're interested in learning more, or find 
more labs for practice at http://openintro.org.
